---
title: "Symulacje Stochastyczne lab 2"
author: "Aleksandra Dąbrowska"
date: "6 March 2017"
output: 
  html_document:
    toc: TRUE
    
---



#Lab1

Chi squared test and p-value.

```{r}

n <- 1000
kostka <- c(1:6)
wyniki <- replicate(n,sample(kostka,1))


chi_2 <- function(x){
  n <- length(x)
  x <- table(x)
  x <- as.data.frame(x)
  x <- x[,2]
  stat <- sum((x-n/6)^2/(n/6))
  p_val <- 1-pchisq(stat, df=5)
  return(c(stat,p_val))
}

chi_2_2 <- function(x){
  n <- length(x)
  stat <- sum((ftable(x)-n/6)^2/(n/6))
  p_val <- 1-pchisq(stat, df=5)
  return(c(stat,p_val))
}


chisq.test(table(wyniki))





```

#Rejection sampling
1.
(A simple Bayesian model) Consider the following 2-stage sampling: first draw θ ∼ Uniform(0, 1) and then X ∼ Bin(n, θ). θ is a parameter of the probability distribution of observed random variable X, and is itself a random variable with the uniform prior distribution. Choose eg. n = 9. • Discover experimentally the marginal distribution of X. Compute this distribution analytically. • Discover experimentally the posterior distribution of θ given X = 3 (say), using ABC. Compute the posterior analytically. Compare the empirical and theoretical distribution. • Try the 2-stage sampling in the opposite order: first draw X from the marginal and then θ from the posterior. Verify that you get θ distributed according to the prior.

a)Discover experimentally the marginal distribution of X.
```{r}


m<- 10000
marginal <- function(){
  theta <- runif(1)
  likelihood <- rbinom(1,10,prob = theta)
  return(likelihood)
}


#wektorem
theta <- runif(10000)
X<- rbinom(10000,10,theta)
hist(X, col="blue")

marginal_distribution <- replicate(m, marginal())
hist(marginal_distribution, col="blue", prob=TRUE)




```


rozkład brzegowy jest jednostajny na zbiorze 11 elementowym {0,...,10}
Prawdopodobieństwo to 1/11=1/n+1


b)Discover experimentally the posterior distribution of θ given X = 3 (say), using ABC
```{r}

abc_alg <- function(){
  X <-0
  while(X!=3){
  theta <- runif(1)
  X<-rbinom(1,10, prob=theta)}
  return(theta)
}

m<- 10000

posterior_dist <- replicate(m,abc_alg())
hist(posterior_dist, col="blue", breaks=200, prob=TRUE)
curve(dnorm(x,4,8), add=TRUE)



```
Rozkład a posteriori to Beta(4,8)


c)Try the 2-stage sampling in the opposite order: first draw X from the marginal and then θ from the posterior. Verify that you get θ distributed according to the prior
```{r}

abc_alg_rev <- function(){
  liczby <- c(0:10)
  X <- sample(liczby,1)
  theta <- rbeta(1,X+1,10-X+1)
  return(theta)
}

a_priori <- replicate(m, abc_alg_rev())
hist(a_priori, prob=TRUE, breaks=200, col="blue")




```

Dostaliśmy rozkład jednostajny na (0,1), czyli rozkład a priori thety.

2.
The following algorithm is a simple example of rejection sampling:
repeat
Gen X ∼ U(0, 1)
Gen U ∼ U(0, 1)
until U < X
return X
What is the distribution of X at the output?
Someone (a rather prominent someone) proposed a ”more efficient version” of this algorithm:
Gen U ∼ U(0, 1)
repeat
Gen X ∼ U(0, 1)
until U < X
return X
What is the distribution of X at the output? Compute theoretically and check the answer
experimentally.

```{r}
m <- 10000


rozkład.X <- function(){
  U <- runif(1)
  X <- 2
  while(U<X){
    X<-runif(1)
  }
  return(X)
}


jaki.x <- replicate(m, rozklad())
hist(jaki.x, prob=TRUE, breaks=100, col="blue")


```

3.Let X1, . . . , Xn, . . . be iid with probability distribution F. Define times and values of records by T1 = 1, R1 = X1, Tk = min{n > Tk−1 : Xn > Rk−1} and Rk = XTk
. Generate a sequence (having the joint distribution equal to that) of record values R1, . . . , Rk for a big k.
• Using brute force, for F = Ex(1) (exponential distribution).
• In a civilized way, for the exponential F.
• For F = Uniform(0, 1), using transformation of the previous sequence.
• Compute 1-dimensional cdf P(Rk ≤ x) theoretically (using induction) and compare with the results of your experiment.


a)
```{r}




```