---
title: "Symulacje stochastyczne 5"
author: "Aleksandra Dąbrowska"
date: "20 March 2017"
output: html_document
---

AR(1).
```{r}
AR <- function(n,sig,ro){
  W <- rnorm(n,0,sig)
  X <- c()
  X[1]<-0
  for (i in 2:n){
    X[i] <- ro*X[i-1]+W[i]
  }
  return(X)
}

n <- 100
ro <- 0.8

wynik_1 <- AR(100,1,0.8)
wynik_2 <- AR(100,1,-0.8)

hist(wynik_1, prob=TRUE)
hist(wynik_2, prob=TRUE)

#plot(1:n, wynik_1, type = "l")
```


Prosty łańcuch Markowa
```{r}
library(markovchain)

alpha <- 0.8
beta <- 0.8

lancuch  <- new("markovchain",states=c("1","2"),transitionMatrix = matrix(data = c(alpha, 1 - alpha, beta, 1 - beta), byrow = TRUE, nrow = 2), name = "stany")

#stany stacjonarne tego lancucha
steadyStates(lancuch)



przebieg <- function(n, alpha, beta) {
      X <- c()
      X[1] <- 1
      for (i in 1:n - 1) {
        X[i+1]<- ifelse(X[i] == 1,  sample(1:2, prob = c(1 - alpha, alpha), size = 1), sample(1:2, prob = c(beta, 1 - beta), size = 1))
      }
      return(X)
}

dane <- przebieg(n,0.9,0.9)

plot(1:n, przebieg(n,0.9,0.9))
